### 2016-09-14 Chapters/SolvationAndCorrelation.tex
CC2 and CC3 written

We will now develop a perturbative many-body expansion of the effective
\acrshort{CC} Lagrangian when a quantum/classical polarizable Hamiltonian is
used.
The use of a \acrshort{CC} ansatz for the wave function ensures term-by-term
size-extensivity of the energy at all orders.~\autocite{Helgaker2000-tz}
The fluctuation potential $\Phi$ is assumed as the perturbation, hence
it will be considered of order 1: $\Phi = O(1)$.
We still assume a closed-shell \acrshort{HF} reference function and
perturbation orders will be counted accordingly.
A superscript index in square brackets - $[i]$ -  will denote an $i$-th
order contribution.
We further assume that real wave functions are used.

The structure of the stationarity conditions,
Eqs.~\eqref{eq:pted-cc-amplitudes} and \eqref{eq:pted-cc-multipliers},
already shows that $\tamp{\mu_u}^{[0]} = 0, \,\forall u$ and
$\tbar{\mu_u}^{[0]} = 0, \, \forall u$.
Thus cluster operators can be of order 1 and higher. We introduce the
following notation for the $i$-th order cluster operators and
multipliers states:
\begin{alignat}{2}
  T_u^{[i]} = \sum_{\mu_u}\tamp{\mu_u}^{[i]}\cluster{\mu_u},
  \quad&
  \bra{\tbar{u}^{[i]}} = \sum_{\mu_u}\tbar{\mu_u}^{[i]}\bra{\mu_u}
\end{alignat}

@@

@@

\todo[inline]{Finish up on CC2 and CC3.}
In complete analogy to the discussion in Section
\ref{sec:cc-approximate}, we can derive iterative and noniterative
schemes for the approximate inclusion of higher order excitations in the
\acrshort*{CC} hierarchy.
Starting from the \acrshort{CC2} and \acrshort{CC3} Lagrangians, Eqs.
\eqref{eq:cc2-lagrangian} and \eqref{eq:cc3-lagrangian} respectively,
we can write down the corresponding effective Lagrangians. The correct
source term $\s_\mathrm{N}(\tamp{},\tbar{})$ must however be identified.
For the \acrshort*{CC2} model, we start from its \acrshort*{CCSD}
expression, Eq. \eqref{eq:source-ccsd}, and retain only terms up to first order:
\begin{equation}
  \begin{aligned}
  \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCSD} &=
  \s_{\mathrm{N}}(\tamp{}, \tbar{})_\mathrm{CCS}
  + \braket{\tbar{1} | \BCHfirst{\s}{T_2} | \mathrm{HF}} \\
  &+ \braket{\tbar{2} | \BCHfirst{\s}{T_2} | \mathrm{HF}}
  + \braket{\tbar{2} | \commutator{\commutator{\s}{T_1}}{T_2} | \mathrm{HF}}
  \end{aligned}
\end{equation}
where the $T_1$-transformed operator was used.

>>>>

We will now develop a perturbative many-body expansion of the effective
\acrshort{PTED}-\acrshort{CC} Lagrangian \eqref{eq:pted-cc}
The use of a \acrshort{CC} ansatz for the wave function ensures term-by-term
size-extensivity of the energy at all orders.~\autocite{Helgaker2000-tz}
The fluctuation potential $\Phi$ is assumed as the perturbation, hence
it will be considered of order 1: $\Phi = O(1)$.
We still assume a closed-shell \acrshort{HF} reference function and
perturbation orders will be counted accordingly.
A superscript index in square brackets - $[i]$ -  will denote an $i$-th
order contribution.
We further assume that real wave functions are used.

The structure of the stationarity conditions,
Eqs. \eqref{eq:pted-cc-amplitudes} and \eqref{eq:pted-cc-multipliers},
already shows that $\tamp{\mu_u}^{[0]} = 0, \,\forall u$ and
$\tbar{\mu_u}^{[0]} = 0, \, \forall u$.
Thus cluster operators can be of order 1 and higher. We introduce the
following notation for the $i$-th order cluster operators and
multipliers states:
\begin{alignat}{2}
  T_u^{[i]} = \sum_{\mu_u}\tamp{\mu_u}^{[i]}\cluster{\mu_u},
  \quad&
  \bra{\tbar{u}^{[i]}} = \sum_{\mu_u}\tbar{\mu_u}^{[i]}\bra{\mu_u}
\end{alignat}

@@

\subsection{Approximations to PTED}

Perturbative expansions can also be developed starting from the
approximations to \acrshort{PTED} presented in Section
\ref{sec:pcm-cc-models}.
For the \acrshort{PTE} scheme, the only change with respect to the
\emph{in vacuo} theory as presented in Section \ref{sec:mbpt} is the use
of the solvated Fock matrix.\autocite{Olivares_del_Valle1991-of,
Angyan1995-co, Lipparini2009-io}

In the \acrshort{PTE(S)} scheme

@@

In complete analogy to the discussion in Section
\ref{sec:cc-approximate}, we can derive iterative and noniterative
schemes for the approximate inclusion of higher order excitations in the
\acrshort*{CC} hierarchy.
Starting from the \acrshort{CC2} Lagrangian, Eq.
\eqref{eq:cc2-lagrangian}, we can write down the corresponding
\acrshort{PTED} effective Lagrangian:\autocite{Sneskov2011-jm,
Schwabe2012-cf}
\begin{equation}\label{eq:pted-cc2-lagrangian}
  \begin{aligned}
  \lag(\tamp{}, \tbar{}, \p)_\mathrm{CC2}
  &=
  E_0
  + \sum_{u=1}^{2}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u}
  + \braket{\mathrm{HF} | \check{\Phi} + \commutator{\check{\Phi}}{Q_2} | \mathrm{HF}} \\
  &+ \braket{\tbar{1} |
  \check{\Phi} + \commutator{\check{\Phi}}{Q_2}
  | \mathrm{HF}}
  + \braket{\tbar{2} | \check{\Phi} | \mathrm{HF}}
  + \frac{1}{2}{}^t\p\V\p +
  {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathrm{CC2}
  \end{aligned}
\end{equation}
where the source term is formed according to Eq.
\eqref{eq:cc2-expectation-value}
The multipliers equations now become:
\begin{subequations}\label{eq:pted-cc2-amplitudes}
  \begin{align}
    \denom{\mu_1}\tamp{\mu_1} &+
    \braket{\mu_1 | \check{\Phi} +
    \commutator{\check{\Phi}}{Q_2} | \mathrm{HF}}
    + {}^t\p\braket{\mu_1 | \check{\s} +
    \commutator{\check{\s}}{Q_2} | \mathrm{HF}}
    = 0
    \label{eq:pted-cc2-singles-amplitudes} \\
    \denom{\mu_2}\tamp{\mu_2} &+ \braket{\mu_2 | \check{\Phi} | \mathrm{HF}}
    + {}^t\p\braket{\mu_2 | \commutator{\check{\s}}{Q_2} | \mathrm{HF}}
    = 0
    \label{eq:pted-cc2-doubles-amplitudes}
  \end{align}
\end{subequations}
while for the multipliers one has:
\begin{subequations}\label{eq:pted-cc2-multipliers}
  \begin{align}
    \denom{\mu_1}\tbar{\mu_1} &+
     \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}}
   + \braket{\tbar{1} |
       \commutator{\check{\Phi}}{\cluster{\mu_1}}
     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{Q_2}
     | \mathrm{HF}} \nonumber \\
   &+ \braket{\tbar{2} | \commutator{\check{\Phi}}{\cluster{\mu_1}} |
   \mathrm{HF}}
   + {}^t\p\braket{\mathrm{HF} |
   \commutator{\check{\s}_\mathrm{N}}{\cluster{\mu_1}} | \mathrm{HF}}  \nonumber \\
   &+ {}^t\p\braket{\tbar{1} | \commutator{\check{\s}}{\cluster{\mu_1}} | \mathrm{HF}}
   + {}^t\p\braket{\tbar{2} | \commutator{\commutator{\check{\s}}{\cluster{\mu_1}}}{Q_2} | \mathrm{HF}}
   = 0 \label{eq:pted-cc2-singles-multipliers} \\
    \denom{\mu_2}\tbar{\mu_2} &+
    \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_2}} | \mathrm{HF}}
   + \braket{\tbar{1} |
       \commutator{\check{\Phi}}{\cluster{\mu_2}}
     | \mathrm{HF}} \nonumber \\
   &+ {}^t\p\braket{\tbar{1} | \commutator{\check{\s}}{\cluster{\mu_2}} | \mathrm{HF}}
   + {}^t\p\braket{\tbar{2} | \commutator{\check{\s}}{\cluster{\mu_2}} | \mathrm{HF}}
     = 0 \label{eq:pted-cc2-doubles-multipliers}
  \end{align}
\end{subequations}
Finally, the polarization is obtained by solving:
\begin{equation}\label{eq:pted-cc2-polarization}
  \V\p + \s_{\mathrm{N}}(\tamp{},\tbar{})_\mathrm{CC2} = 0
\end{equation}

The \acrshort{PTED}-\acrshort{CC3} model is obtained similarly. The
source term is formed according to Eq. \eqref{eq:cc3-expectation-value}.
Thus, adding the polarization functional to the Lagrangian
\eqref{eq:cc3-lagrangian}, we obtain the effective Lagrangian:
\begin{equation}\label{eq:pted-cc3-lagrangian}
  \begin{aligned}
  \lag(\tamp{}, \tbar{}, \p)_\mathrm{CC3}
  E_0
  + \sum_{u=1}^{3}\tbar{\mu_u}\denom{\mu_u}\tamp{\mu_u}
  + \braket{\mathrm{HF} | \check{\Phi} + \commutator{\check{\Phi}}{T_2} | \mathrm{HF}} \\
  &+ \braket{\tbar{1} |
  \check{\Phi} + \commutator{\check{\Phi}}{T_2} + \commutator{\check{\Phi}}{Q_3}
  | \mathrm{HF}} \\
  &+ \braket{\tbar{2} |
  \check{\Phi} + \commutator{\check{\Phi}}{T_2}
  + \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2} + \commutator{\check{\Phi}}{Q_3}
  | \mathrm{HF}} \\
  &+ \braket{\tbar{3} | \commutator{\check{\Phi}}{T_2} | \mathrm{HF}}
  + \frac{1}{2}{}^t\p\V\p +
  {}^t\p\s_{\mathrm{N}}(\tamp{},\tbar{})_\mathrm{CC3}
  \end{aligned}
The polarization equation is straightforwardly:
\begin{equation}\label{eq:pted-cc3-polarization}
  \V\p + \s_{\mathrm{N}}(\tamp{},\tbar{})_\mathrm{CC3} = 0
The amplitudes are determined by:
\begin{subequations}\label{eq:pted-cc3-amplitudes}
    \denom{\mu_1}\tamp{\mu_1} &+ \braket{\mu_1 |
    \check{\Phi} + \commutator{\check{\Phi}}{T_2}
    + \commutator{\check{\Phi}}{Q_3}
    | \mathrm{HF}} \nonumber \\
    &+ {}^t\p\braket{\mu_1 | \check{\s} +
    \commutator{\check{\s}}{T_2} | \mathrm{HF}}
    = 0 \label{eq:pted-cc3-singles-amplitudes} \\
    \denom{\mu_2}\tamp{\mu_2} &+ \braket{\mu_2 |
    \check{\Phi} + \commutator{\check{\Phi}}{T_2}
    + \frac{1}{2}\commutator{\commutator{\check{\Phi}}{T_2}}{T_2}
    + \commutator{\check{\Phi}}{Q_3}
    | \mathrm{HF}} \nonumber  \\
    {}^t\p\braket{\mu_2 | \commutator{\check{\s}}{T_2} +
    \commutator{\check{\s}}{Q_3} | \mathrm{HF}}
    = 0 \label{eq:pted-cc3-doubles-amplitudes} \\
    \denom{\mu_3}\tamp{\mu_3} &+ \braket{\mu_3 |
    \commutator{\check{\Phi}}{T_2}
    | \mathrm{HF}}
    + {}^t\p\braket{\mu_3 | \commutator{\check{\s}}{Q_3}
     + \frac{1}{2}\commutator{\commutator{\check{\s}}{T_2}}{T_2}
    | \mathrm{HF}}
    = 0 \label{eq:pted-cc3-triples-amplitudes}
and finally the multipliers obey:
\begin{subequations}\label{eq:pted-cc3-multipliers}
    \denom{\mu_1}\tbar{\mu_1} &+
     \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_1}} | \mathrm{HF}}
   + \braket{\tbar{1} |
       \commutator{\check{\Phi}}{\cluster{\mu_1}}
     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}
     | \mathrm{HF}} \nonumber \\
   &+ \braket{\tbar{2} |
       \commutator{\check{\Phi}}{\cluster{\mu_1}}
     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}
     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{Q_3}
   | \mathrm{HF}} \nonumber \\
   &+ \braket{\tbar{3} |
      \commutator{\commutator{\check{\Phi}}{\cluster{\mu_1}}}{T_2}
     | \mathrm{HF}}
   + {}^t\p\braket{\mathrm{HF} |
   \commutator{\check{\s}_\mathrm{N}}{\cluster{\mu_1}} | \mathrm{HF}}  \nonumber \\
   &+ {}^t\p\braket{\tbar{1} | \commutator{\check{\s}}{\cluster{\mu_1}} | \mathrm{HF}}
   + {}^t\p\braket{\tbar{2} |
   \commutator{\commutator{\check{\s}}{\cluster{\mu_1}}}{Q_2} |
   \mathrm{HF}} \nonumber \\
   &+ {}^t\p\braket{\tbar{3} | \commutator{\commutator{\check{\s}}{\cluster{\mu_1}}}{Q_3} | \mathrm{HF}}
     = 0 \label{eq:pted-cc3-singles-multipliers} \\
    \denom{\mu_2}\tbar{\mu_2} &+
    \braket{\mathrm{HF} | \commutator{\check{\Phi}}{\cluster{\mu_2}} | \mathrm{HF}}
   + \braket{\tbar{1} |
       \commutator{\check{\Phi}}{\cluster{\mu_2}}
     | \mathrm{HF}} \nonumber \\
  &+ \braket{\tbar{2} |
       \commutator{\check{\Phi}}{\cluster{\mu_2}}
     + \commutator{\commutator{\check{\Phi}}{\cluster{\mu_2}}}{T_2}
     | \mathrm{HF}}
  + \braket{\tbar{3} |
       \commutator{\check{\Phi}}{\cluster{\mu_2}}
     | \mathrm{HF}} \nonumber \\
   &+ {}^t\p\braket{\tbar{1} | \commutator{\check{\s}}{\cluster{\mu_2}} | \mathrm{HF}}
   + {}^t\p\braket{\tbar{2} | \commutator{\check{\s}}{\cluster{\mu_2}} | \mathrm{HF}}
   &+ {}^t\p\braket{\tbar{3} |
   \commutator{\commutator{\check{\s}}{\cluster{\mu_2}}}{T_2} | \mathrm{HF}}
    = 0 \label{eq:pted-cc3-doubles-multipliers} \\
    \denom{\mu_3}\tbar{\mu_3} &+
    \braket{\tbar{1} |
       \commutator{\check{\Phi}}{\cluster{\mu_3}}
     | \mathrm{HF}}
  + \braket{\tbar{2} |
       \commutator{\check{\Phi}}{\cluster{\mu_3}}
     | \mathrm{HF}} \nonumber \\
   &+ {}^t\p\braket{\tbar{2} | \commutator{\check{\s}}{\cluster{\mu_3}} | \mathrm{HF}}
   + {}^t\p\braket{\tbar{3} | \commutator{\check{\s}}{\cluster{\mu_3}} | \mathrm{HF}}
    = 0 \label{eq:pted-cc3-triples-multipliers}
The \acrshort{PTES}, \acrshort{PTE(S)} and \acrshort{PTE} approximate
schemes can be applied, \emph{mutatis mutandis}, also within the
\acrshort{CC2} and \acrshort{CC3} models.\autocite{Caricato2011-tx, Schwabe2012-cf, Krause2016-ee}
In particular, we note that the \acrshort{PTES} approximation is crucial
in the implementation of \acrshort{PE}-\acrshort{CC2} presented by
\citeauthor{Schwabe2012-cf} in \noparcite[ref.][]{Schwabe2012-cf}
The resolution of the identity can in fact still be applied when
assuming the \acrshort{PTES} approximation.
Moreover, the approximate \acrshort{PE}-\acrshort{CC} Lagrangian of
\citeauthor{Krause2016-ee} is essentially the \acrshort{PTE(S)}
approximation for \acrshort{PE}-\acrshort{CC}.\autocite{Krause2016-ee}
